Subglacial and subaerial fluvial sediment transport capacity respond differently to water discharge variations

### 2024-07-04 Delaneyetal2024.tex
Updates from Overleaf
change on machine

Despite the physical differences between the subaerial and subglacial systems, water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
Many of these studies discuss the disparity between water discharge and sediment export, highlighting the non-linear relationship between the two, along with the role of sediment access or production.
Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited, or hydraulically controlled, regime.
The response of sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in these glacierized catchment systems \citep[e.g.][]{ganti2016,mancini2023}.
This relationship must be clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.

@@

Across the range of parameters examined (Section\,\ref{sect:imp}), variability in all model outputs remains higher in the subglacial system compared to the subaerial one for both \alpine{} and \icesheet{} (Figure~\ref{fig:multi_run}).
In some cases, subglacial model outputs' variability is a magnitude larger than their subaerial counterparts.
Results show that variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately $1$--$5$ days (Figure~\ref{fig:multi_run}).
These smoothing timescales remove the diurnal variations in water discharge.
In the \icesheet case, the smaller variability in model outputs compared to \alpine means that the R-channel is in closer equilibrium between channel size and water discharge (Section~\ref{sect:scaling}).
In this case, water discharge variations would result in smaller sediment transport capacity variations than even in subaerial cases as the subglacial channel size evolves to accommodate them (Table~\ref{tab:Qs}).

>>>>

Despite the physical differences between the subaerial and subglacial systems, water discharge and sediment export are often compared in contemporary glacierized catchments \citep[e.g.][]{willis1996,hodson1998,pearce2003,richards2003,swift2005,chu2009,tedstone2012,chu2012,overeem2017,delaney2018,swift2021,lu2022,andresen2024}.
Many of these studies discuss the disparity between water discharge and sediment export, particularly the non-linear relationship between the two, along with the role of sediment access or production.
Yet, a variable relationship between water discharge and sediment export could be expected from glaciers with evolving conduit size in a transport-limited, or hydraulically controlled, regime.
The response of sediment transport capacity in subglacial sediment transport may affect the interpretation of sediment transport records in these glacierized catchment systems \citep[e.g.][]{ganti2016,mancini2023}.
This relationship must be clarified as sediment dynamics change in glacierized regions along with changing hydrology \citep[e.g.][]{brunner2019}, increasing variability of water discharge \citep{lane2019} and occurrence of extreme glacier melt  \citep[e.g.][]{overeem2015,cremona2023} that may drive high magnitude sediment transport events.

@@

Across the range of parameters examined (Section\,\ref{sect:imp}), variability in all model outputs remains higher in the subglacial system compared to the subaerial one for both \alpine{} and \icesheet{} (Figure~\ref{fig:multi_run}).
In some cases, subglacial model outputs' variability is a magnitude larger than their subaerial counterparts.
Results show that variability in both subglacial and subaerial outputs decreases with smoothing time longer than approximately $1$--$5$ days (Figure~\ref{fig:multi_run}).
These smoothing timescales remove the diurnal variations in water discharge.
